---
title: "Getting Compliance Documentation"
output: html_document
---

**Author**: Thodoris Petropoulos

**Label**: Model Management
### Scope
The scope of this notebook is to provide instructions on how to get Compliance Documentation documents using the R API.

### Background
Compliance documentation is a premium add-on product to DataRobot. It allows users to automatically generate and download documentation to assist with deploying models in highly regulated industries.

### Requirements

- R version 3.6.2
-  DataRobot API version 2.16.0. 
Small adjustments might be needed depending on the R version and DataRobot API version you are using.

Full documentation of the R package can be found here: https://cran.r-project.org/web/packages/datarobot/index.html

It is assumed you already have a DataRobot <code>Project</code> object and a DataRobot <code>Model </code> object.

#### Import Packages

```{r results = 'hide', warning=FALSE, message=FALSE}
library(datarobot)
```

#### Downloading Compliance Documentation
To download compliance documentation for a particular model, call <code>DownloadComplianceDocumentation</code> on a particular model and specify a filepath to download the documentation to. Note that it downloads in DOCX format.

```{r eval=FALSE}
DownloadComplianceDocumentation(model, "path/to/filename.docx")
```

### Creating a Custom Template
You can also use your own custom compliance documentation templates.

#### The Default Template
First, let’s get the default template. This can be done just by using <code>GetComplianceDocTemplate</code> It downloads as a JSON file.

```{r eval=FALSE}
GetComplianceDocTemplate("path/to/filename.json")
```

#### Updating the Default Template
A common workflow for building your own template is downloading the default template and modifying it.

```{r eval=FALSE}
DownloadComplianceDocTemplate("path/to/filename.json")
# ...then modify the compliance doc template in your favorite editor.
UploadComplianceDocTemplate(name = "myNewTemplate", filename = "path/to/modified_file.json")
```

Alternatively, you can construct a template via a list:

```{r eval=FALSE}
sections <- list(list("title" = "Missing Values Report",
                      "highlightedText" = "NOTICE",
                      "regularText" = "This dataset had a lot of Missing Values. See the chart below: {{missingValues}}",
                      "type" = "user"),
                 list("title" = "Blueprints",
                      "regularText" = "{{blueprintDiagram}} /n Blueprint for this model",
                      "type" = "user"))
UploadComplianceDocTemplate(name = "myNewTemplateFromSections", sections = sections)
```

You can then get and download your template:

```{r eval=FALSE}
myTemplate <- ListComplianceDocTemplates(namePart = "myNewTemplateFromSections")[[1]]
DownloadComplianceDocTemplate(myTemplate)

```