---
title: "Starting a Binary Classification Project"
output: html_document
---

**Author**: Thodoris Petropoulos

**Label**: Modeling Options

### Scope
The scope of this notebook is to provide instructions on how to initiate a DataRobot project manually where the user has the option to choose which models/blueprints he wants to initiate. The procedure below should work for any type of problem you are trying to solve (regression, classification, time series, etc).

### Requirements

- R version 3.6.2
-  DataRobot API version 2.16.0. 
Small adjustments might be needed depending on the R version and DataRobot API version you are using.

Full documentation of the R package can be found here: https://cran.r-project.org/web/packages/datarobot/index.html

#### Import Packages

```{r results = 'hide', warning=FALSE, message=FALSE}
library(datarobot)
library(mlbench)
```

#### Import Dataset
We will be loading the breast cancer dataset. A very simple binary classification dataset that is available through the mlbench package.

```{r}
data(BreastCancer)
head(BreastCancer)
```

#### Connect to DataRobot
Connect to DataRobot using your credentials and your endpoint. Change input below accordingly.
```{r results = 'hide', warning=FALSE, message=FALSE}
ConnectToDataRobot(endpoint = "YOUR_HOSTNAME", 
                   token = "YOUR_API_KEY")
```

#### Initiate Project
We will be initiating the project using <code>autopilot_on = False</code>. This way DataRobot will not start building models until we specify which ones we want to build

```{r}
project <- StartProject(dataSource = BreastCancer,
                        projectName = "myBinaryClassificationProject",
                        target = "Class",
                        mode = "manual"
                        )
```

#### Find all of the blueprints
We can use the <code>Listblueprints</code> method to see all of the blueprints DataRobot generated.

```{r}
blueprints <- ListBlueprints(project)
```

Now that we have the Blueprints, we can search for a specific blueprint.
For example, all models that have "Gradient" in their name.

```{r}
models_to_run<- blueprints[grepl("Gradient", unlist(lapply(blueprints,"[", "modelType")))]
```

#### Lets now initiate these models
We can use the <code>RequestNewModel</code> method to initiate modeling for a specific blueprint.

```{r}
for (i in 1:length(models_to_run)){
  job <- RequestNewModel(project, models_to_run[[i]])
  WaitForJobToComplete(project, job,maxWait=600)
}
```
