---
title: "Starting a Regression Project"
output: html_document
---

**Author**: Thodoris Petropoulos

**Label**: Modeling Options

### Scope
The scope of this notebook is to provide instructions on how to initiate a DataRobot project for a numerical target using the R API.

### Background
Regression Analysis is the task of predicting the value of a continuous target column.

Examples:

- Predict Life Time Value (LTV) of customer.
- Predicting player performance.
- Predicting house price.

The target column will always be a continuous numeric variable even though regression could also be applicable a discreet high cardinality variable.

### Requirements

- R version 3.6.2
-  DataRobot API version 2.16.0. 
Small adjustments might be needed depending on the R version and DataRobot API version you are using.

Full documentation of the R package can be found here: https://cran.r-project.org/web/packages/datarobot/index.html


#### Import Packages

```{r results = 'hide', warning=FALSE, message=FALSE}
library(datarobot)
library(mlbench)

```

#### Import Dataset
We will be loading the Boston Housing dataset. A very simple dataset for regression that is available through mlbench.

```{r}
data(BostonHousing)
head(BostonHousing)
```
#### Connect to DataRobot
Connect to DataRobot using your credentials and your endpoint. Change input below accordingly.
```{r results = 'hide', warning=FALSE, message=FALSE}
ConnectToDataRobot(endpoint = "YOUR_DATAROBOT_HOSTNAME", 
                   token = "YOUR_API_KEY")
```

#### Initiate Project
I will be initiating a project using <code>StartProject </code>:

* dataSource: Data source (Could be path to file or R dataframe)
* projectName: Name of Project
* target: String with target variable name
* workerCount: Amount of workers to use
* metric: Optimisation metric to use
* wait: Waits for autopilot to complete before moving to next chunk of code 

```{r eval=FALSE}
project <- StartProject(dataSource = BostonHousing,
                        projectName = "MyRegressionProject",
                        target = "medv",
                        wait = TRUE) #Wait equals True means wait for autopilot to finish
```