---
title: "Starting a Binary Classification Project"
output: html_document
---

**Author**: Thodoris Petropoulos

**Label**: Modeling Options

### Scope
The scope of this notebook is to provide instructions on how to initiate a DataRobot project for a Binary Classification target using the R API.

### Background
Binary classification is the task of classifying the elements of a given set into two groups.

Examples:

- A customer is a churner or not.
- A loan is going to default or not.
- A patient has a disease or not.

Most commonly, the target column will have values:

- 0/1
- Yes/No
- True/False

### Requirements

- R version 3.6.2
-  DataRobot API version 2.16.0. 
Small adjustments might be needed depending on the R version and DataRobot API version you are using.

Full documentation of the R package can be found here: https://cran.r-project.org/web/packages/datarobot/index.html

#### Import Packages

```{r results = 'hide', warning=FALSE, message=FALSE}
library(datarobot)
library(mlbench)
```

#### Import Dataset
We will be loading the breast cancer dataset. A very simple binary classification dataset that is available through the mlbench package.

```{r}
data(BreastCancer)
head(BreastCancer)
```

#### Connect to DataRobot
Connect to DataRobot using your credentials and your endpoint. Change input below accordingly.
```{r results = 'hide', warning=FALSE, message=FALSE}
ConnectToDataRobot(endpoint = "YOUR_DATAROBOT_HOSTNAME", 
                   token = "YOUR_API_KEY")
```

#### Initiate Project
I will be initiating a project using <code>StartProject </code>:

* dataSource: Data source (Could be path to file or R dataframe)
* projectName: Name of Project
* target: String with target variable name
* workerCount: Amount of workers to use
* metric: Optimisation metric to use
* wait: Waits for autopilot to complete before moving to next chunk of code 

```{r eval=FALSE}
project <- StartProject(dataSource = BreastCancer,
                        projectName = "myBinaryClassificationProject",
                        target = "Class",
                        wait = TRUE)
```