---
title: "Feature Lists Manipulation"
output: html_document
---

**Author**: Thodoris Petropoulos

**Label**: Modeling Options

### Scope
The scope of this notebook is to provide instructions on how to create and manipulate custom feature lists using the R API as well as how to train a model using a custom feature list.

### Background
Feature lists control the subset of features that DataRobot uses to build models. 

### Requirements

- R version 3.6.2
-  DataRobot API version 2.16.0. 
Small adjustments might be needed depending on the R version and DataRobot API version you are using.

Full documentation of the R package can be found here: https://cran.r-project.org/web/packages/datarobot/index.html

It is assumed you already have a DataRobot <code>Project</code> object and a DataRobot <code>Model </code> object.

#### Import Packages
```{r results = 'hide', warning=FALSE, message=FALSE}
library(datarobot)
```

#### Listing Available Feature Lists

```{r echo=FALSE, results = 'hide', warning=FALSE, message=FALSE}
#This piece of code will not show
ConnectToDataRobot(endpoint = "YOUR_DATAROBOT_HOSTNAME", 
                   token = "YOUR_API_KEY")

project <- GetProject("YOUR_PROJECT_ID")
allModels <- ListModels(project)
modelFrame <- as.data.frame(allModels)
metric <- modelFrame$validationMetric
if (project$metric %in% c('AUC', 'Gini Norm')) {
  bestIndex <- which.max(metric)
} else {
  bestIndex <- which.min(metric)
}
model <- allModels[[bestIndex]]
blueprint = model$blueprint
```

#### Listing Available Feature Lists

```{r}
ListFeaturelists(project)
```

#### Creating a custom Feature List
To create a custom feature list use the <code>CreateFeaturelist</code> command.

```{r}
custom_featurelist <- CreateFeaturelist(project, 'CustomFeaturelist', c('race','age','gender'))
```

#### Training a model with a custom feature list
```{r eval=FALSE}
blueprint <- GetBlueprint(project, model$blueprintId)
RequestNewModel(project,blueprint=blueprint, featurelist=custom_featurelist)
```