---
title: "Prediction Explanation Clustering with R"
output: html_notebook
---

This notebook is intended to accompany the [*Prediction Explanation Clustering with R* post](https://community.datarobot.com/t5/resources/prediction-explanation-clustering-with-r/ta-p/7862) on the DataRobot Community. The post, together with the code below, illustrates how to use the `datarobot.pe.clustering` R package, available at the [pe-clustering-R repository](https://github.com/datarobot-community/pe-clustering-R) on the DataRobot Community GitHub.

# Installing the Package

```{r}
if (!require("remotes")) { install.packages("remotes") }
if (!require("datarobot.pe.clustering")) { remotes::install_github("datarobot-community/pe-clustering-R", build_vignettes=TRUE, upgrade = "never") }
```

# Loading libraries

```{r}
library(datarobot.pe.clustering)
library(ggplot2)
library(dplyr)
library(tidyr)
```


# Setting up the data

```{r}
library(mlbench)
data(PimaIndiansDiabetes)
head(PimaIndiansDiabetes)
```


# Obtaining a DataRobot model

```{r}
project <- StartProject(dataSource = PimaIndiansDiabetes,
                        projectName = "PredictionExplanationClusteringVignette",
                        target = "diabetes",
                        mode = "quick",
                        wait = TRUE)
models <- ListModels(project$projectId)
model <- models[[1]]
```

```{r}
summary(model)['modelType']
```


# Running prediction explanations

```{r}
scoring_df <- PimaIndiansDiabetes %>% select(-diabetes)
```

```{r}
results <- cluster_and_summarize_prediction_explanations(
      model,
      scoring_df,
      num_feature_summarizations=10,
      num_neighbors=50,
      min_dist=10^-100,
      min_points=25
    )
```

```{r}
str(results, max.level = 1)
```


# Introspection of results

## Summary and Plot

```{r}
summary(results)
```

```{r}
plot(results)
```

```{r}
ggplot(results$plot_data, aes(x=dim1, y=dim2, color=clusterID)) +
  geom_point()+
  theme_bw()+
  labs(title='Records by Prediction Explanation Cluster', x='Reduced Dimension 1', y='Reduced Dimension 2')
```


## Characterizing clusters by prediction risk and feature values

```{r}
scoring_df_with_clusters <- scoring_df
scoring_df_with_clusters$cluster <- factor(results$cluster_ids)
scoring_df_with_clusters$predicted_risk <- results$pe_frame$class1Probability
```

```{r}
scoring_df_with_clusters %>%
    ggplot(aes(x=cluster,y=predicted_risk, fill=cluster))+geom_violin()+
    labs(title='Predicted Diabetes Risk by Cluster')+
    theme_bw()
```

```{r}
scoring_df_with_clusters %>%
    gather(key='feature',value='value',-cluster)%>%
    ggplot(aes(x=value, group=cluster, color=cluster, fill=cluster)) +
    geom_density(alpha=0.2)+
    facet_wrap(~feature, scales='free')+
    theme_bw()
```


## Characterizing clusters by prediction explanation strength

```{r}
strength_matrix_with_clusters <- results$strength_matrix
strength_matrix_with_clusters$cluster <- factor(results$cluster_ids)
head(strength_matrix_with_clusters)
```

```{r}
strength_matrix_with_clusters %>%
  gather(feature, strength, -cluster)%>%
  ggplot(aes(x=strength, group=cluster, color=cluster, fill=cluster)) +
  geom_density(alpha=0.2)+
  facet_wrap(~feature, scales='free')+
  xlab('Strength of prediction explanation')+
  theme_bw()
```

