---
title: "Getting_Lift_Chart"
output: html_document
---

**Author**: Peter Hurford

**Label**: Evaluating Models

### Scope
The scope of this notebook is to provide instructions on how to get Lift Chart of a specific model using the R API.


### Background
Insights provided by the Lift Chart are helpful in checking the performance of machine learning models. 

The lift chart depicts how well a model segments the target population and how capable it is of predicting the target across the prediction distribution.  This allows you to evaluate the performance of a model at both the high end (very likely) and low end (very unlikely) end of the spectrum. 

For example, so you can see how well the model performs for different ranges of values of the target variable. Looking at the Lift chart, the left side of the curve indicates where the model predicted a low score on one section of the population while the right side of the curve indicates where the model predicted a high score. In general, the steeper the actual line is, and the more closely the predicted line matches the actual line, the better the model is. A consistently increasing line is another good indicator.

### Requirements

- R version 3.6.2
-  DataRobot API version 2.16.0. 
Small adjustments might be needed depending on the R version and DataRobot API version you are using.

Full documentation of the R package can be found here: https://cran.r-project.org/web/packages/datarobot/index.html

It is assumed you already have a DataRobot <code>Project</code> object and a DataRobot <code>Model</code> object.

#### Import Packages

```{r results = 'hide', warning=FALSE, message=FALSE}
library(datarobot)
library(ggplot2)
```

#### Requesting Lift Chart Data

Lift chart data can be retrieved for a specific data partition (validation, cross-validation, or holdout) or for all the data partitions using <code>GetLiftChart</code> and <code>ListLiftCharts</code>. To retrieve the data for holdout partition, it needs to be unlocked first.

Let’s retrieve the validation partition data for top model using <code>GetLiftChart</code>. The <code>GetLiftChart</code> function returns data for validation partition by default. We can retrieve data for specific data partition by passing value to source parameter in <code>GetLiftChart</code>.

```{r echo=FALSE, results = 'hide', warning=FALSE, message=FALSE}
#This piece of code will not show
ConnectToDataRobot(endpoint = "YOUR_DATAROBOT_HOSTNAME", 
                   token = "YOUR_API_KEY")

project <- GetProject("YOUR_PROJECT_ID")
allModels <- ListModels(project)
saveRDS(allModels, "modelsModelInsights.rds")
modelFrame <- as.data.frame(allModels)
metric <- modelFrame$validationMetric
if (project$metric %in% c('AUC', 'Gini Norm')) {
  bestIndex <- which.max(metric)
} else {
  bestIndex <- which.min(metric)
}
model <- allModels[[bestIndex]]
model$modelType
```


```{r}
lc <- GetLiftChart(model)
saveRDS(lc, "liftChartModelInsights.rds")
head(lc)
```

#### Plotting Lift Chart
```{r}
lc$actual <- lc$actual / lc$binWeight
lc$predicted <- lc$predicted / lc$binWeight
lc <- lc[order(lc$predicted), ]
lc$binWeight <- NULL
lc <- data.frame(value = c(lc$actual, lc$predicted),
                 variable = c(rep("Actual", length(lc$actual)),
                              rep("Predicted", length(lc$predicted))),
                 id = rep(seq_along(lc$actual), 2))
ggplot(lc) + geom_line(aes(x = id, y = value, color = variable))
```

