---
title: "Getting Feature Impact"
output: html_document
---

**Author**: Peter Hurford

**Label**: Evaluating Models

### Scope
The scope of this notebook is to provide instructions on how to get Feature Impact scores of a specific model using the R API.

### Background
Feature Impact informs us of how important a feature is in a model agnostic way. Impact is normalised so most impactful feature will always have a a feature impact score of 100%.

### Requirements

- R version 3.6.2
-  DataRobot API version 2.16.0. 
Small adjustments might be needed depending on the R version and DataRobot API version you are using.

Full documentation of the R package can be found here: https://cran.r-project.org/web/packages/datarobot/index.html

It is assumed you already have a DataRobot <code>Project</code> object and a DataRobot <code>Model</code> object.

#### Import Packages
```{r results = 'hide', warning=FALSE, message=FALSE}
library(datarobot)
library(ggplot2)
```

#### Requesting Feature Impact

```{r echo=FALSE, results = 'hide', warning=FALSE, message=FALSE}
#This piece of code will not show
ConnectToDataRobot(endpoint = "YOUR_DATAROBOT_HOSTNAME", 
                   token = "YOUR_API_KEY")

project <- GetProject("YOUR_PROJECT_ID")
allModels <- ListModels(project)
modelFrame <- as.data.frame(allModels)
metric <- modelFrame$validationMetric
if (project$metric %in% c('AUC', 'Gini Norm')) {
  bestIndex <- which.max(metric)
} else {
  bestIndex <- which.min(metric)
}
model <- allModels[[bestIndex]]
model$modelType
```

```{r}
#Get Feature Impact
feature_impact = GetFeatureImpact(model)
head(feature_impact)
```

#### Plotting Feature Impact
```{r}
ggplot(head(feature_impact,5), aes(x=featureName, y=impactNormalized)) + 
  geom_bar(stat = "identity")
```