---
title: "Getting_ROC_Curve"
output: html_document
---

**Author**: Peter Hurford

**Label**: Evaluating Models

### Scope
The scope of this notebook is to provide instructions on how to get ROC Curve data of a specific model using the R API.

### Background
Insights provided by the ROC Curve are helpful in checking the performance of machine learning models. 

The receiver operating characteristic curve, or ROC curve, is a graphical plot that illustrates the performance of a binary classifier system as its discrimination threshold is varied. The curve is created by plotting the true positive rate (TPR) against the false positive rate (FPR) at various threshold settings.

### Requirements

- R version 3.6.2
-  DataRobot API version 2.16.0. 
Small adjustments might be needed depending on the R version and DataRobot API version you are using.

Full documentation of the R package can be found here: https://cran.r-project.org/web/packages/datarobot/index.html

It is assumed you already have a DataRobot <code>Project</code> object and a DataRobot <code>Model</code> object.

#### Import Packages

```{r results = 'hide', warning=FALSE, message=FALSE}
library(datarobot)
library(ggplot2)
```
#### Requesting ROC Curve Data

ROC curve data can be generated for a specific data partition (validation, cross validation, or holdout) or for all the data partition using the <code>GetRocCurve</code> method.

To retrieve ROC curve information:

```{r echo=FALSE, results = 'hide', warning=FALSE, message=FALSE}
#This piece of code will not show
ConnectToDataRobot(endpoint = "YOUR_DATAROBOT_HOSTNAME", 
                   token = "YOUR_API_KEY")

project <- GetProject("YOUR_PROJECT_ID")
allModels <- ListModels(project)
modelFrame <- as.data.frame(allModels)
metric <- modelFrame$validationMetric
if (project$metric %in% c('AUC', 'Gini Norm')) {
  bestIndex <- which.max(metric)
} else {
  bestIndex <- which.min(metric)
}
model <- allModels[[bestIndex]]
model$modelType
```

```{r}
roc <- GetRocCurve(model)
head(roc)
```

#### Plotting ROC Curve
```{r}
ValidationRocCurve <- GetRocCurve(model)
ValidationRocPoints <- ValidationRocCurve[["rocPoints"]]
ggplot(ValidationRocPoints, aes(x = falsePositiveRate, y = truePositiveRate)) + geom_line()
```

#### Threshold Operations
You can get the recommended threshold value with maximal F1 score. That is the same threshold that is preselected in DataRobot when you open the “ROC curve” tab.

```{r}
threshold <- ValidationRocPoints$threshold[which.max(ValidationRocPoints$f1Score)]
```

You can also estimate metrics for different threshold values. This will produce the same results as updating the threshold on the DataRobot “ROC curve” tab.

```{r}
ValidationRocPoints[ValidationRocPoints$threshold == tail(Filter(function(x) x > threshold,
                                                                 ValidationRocPoints$threshold), 1), ]
```