---
title: "Getting_Word_Cloud"
output: html_document
---

**Author**: Peter Hurford

**Label**: Evaluating Models

### Scope
The scope of this notebook is to provide instructions on how to get the Word Cloud of a specific model using the R API.

### Background
The word cloud is a type of insight available for some text-processing models for datasets containing text columns. You can get information about how the appearance of each ngram (word or sequence of words) in the text field affects the predicted target value.

This example will show you how to obtain word cloud data and visualize it, similar to how DataRobot visualizes the word cloud in the “Model Insights” tab interface.

### Requirements

- R version 3.6.2
-  DataRobot API version 2.16.0. 
Small adjustments might be needed depending on the R version and DataRobot API version you are using.

Full documentation of the R package can be found here: https://cran.r-project.org/web/packages/datarobot/index.html

It is assumed you already have a DataRobot <code>Project</code> object and a DataRobot <code>Model </code> object.

#### Import Packages
```{r results = 'hide', warning=FALSE, message=FALSE}
library(datarobot)
library(colormap)
library(modelwordcloud)
```

```{r echo=FALSE, results = 'hide', warning=FALSE, message=FALSE}
#This piece of code will not show
ConnectToDataRobot(endpoint = "YOUR_DATAROBOT_HOSTNAME", 
                   token = "YOUR_API_KEY")

project <- GetProject("YOUR_PROJECT_ID")
project$projectName
allModels <- ListModels(project)
```

#### Requesting Word Cloud
Request word cloud with <code>GetWordCloud</code> function.

```{r}
 allModels2 <- as.data.frame(allModels, simple = TRUE)

# Find first available model with word-based capabilities
for(i in 1:length(allModels2$modelId))
{
modelId <- allModels2$modelId[i]
model <- GetModel(project$projectId, modelId)
if(GetModelCapabilities(model)$hasWordCloud == TRUE)
  {
  # Get word cloud
  wordCloud <- GetWordCloud(project, model$modelId)
  }
}
```

#### Plotting Word Cloud
```{r}
# Remove stop words
wordCloud <- wordCloud[!wordCloud$isStopword, ]

# Make word cloud
colors <- c(colormap(c("#255FEC", "#2DBEF9")), colormap(c("#FFAC9D", "#D80909"), reverse = TRUE))
suppressWarnings(wordcloud(words = wordCloud$ngram,
                           freq = wordCloud$frequency,
                           coefficients = wordCloud$coefficient,
                           colors = colors,
                           scale = c(3, 0.3)))
```